Hydras \& Co.

### 2022-04-12 doc/gaia-chapter.tex
corrections in gaia chapter

\subsubsection{Remark}
\label{sect:gaia-masking}
We try to respect the following ``protocol''.
Let us assume we adapt some definitions and properties
from a given module \texttt{hydras.Epsilon0.Foo}, and a constant \texttt{bar} of this module, whose absolute name is
thus \texttt{hydras.Epsilon0.Foo.bar}.
\begin{enumerate}
\item We create a module \texttt{gaia\_hydras.GFoo}
  (``G'' for \gaia).
\item Inside \texttt{GFoo}, the identifier \texttt{bar}
  is reserved in priority to \gaia's version, and the notation
  \texttt{hbar} is used for \texttt{hydras.Epsilon0.Foo.bar}.
\end{enumerate}

@@

Natural sum  (a.k.a. Hessenberg's  sum) is defined in
\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).

We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.

\vspace{4pt} 
\noindent.

@@

\inputsnippets{GHessenberg/oplusEquations}
\inputsnippets{GHessenberg/oplusEquationsb}
\inputsnippets{GHessenberg/oplusEquationsc}

\inputsnippets{GHessenberg/oplusNf}

@@

First, we rename \HydrasLib' \texttt{canon} function into
\texttt{hcanon}. Then we define a new function \emph{via} the
translations \texttt{g2h} and \texttt{h2g}.

\inputsnippets{GCanon/canonDef}

\inputsnippets{GCanon/g2hCanonRw}

@@

\inputsnippets{GCanon/gcanonLt}
\inputsnippets{GCanon/gcanonLimitMono}
\inputsnippets{GCanon/glimitCanonSNotZero}

\subsubsection{Canonical sequences and limits}

@@

We adapt theorems from \href{../theories/html/hydras.Epsilon0.Canon.html}%
{\texttt{hydras.Epsilon0.Canon}} to \gaia's vocabulary.

\inputsnippets{GCanon/gcanonLimitStrong}
\inputsnippets{GCanon/gcanonLimitV2}
\inputsnippets{GCanon/gcanonLimitOf}

@@

@@

\section{Rapidly growing arithmetic functions}
In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$. In \HydrasLib, these functions are defined by transfinite induction, using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}. For that purpose, we defined a sigma-type
\texttt{hydras.Epsilon0.E0.E0} of ordinal-terms in Cantor normal form.

These families share some monotony properties, also re-defined in
\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge}:

\inputsnippets{T1Bridge/MonoDef}
\subsection{A type for well-formed ordinal terms}
\label{sect:E0-gaia-hydras}
For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.

\inputsnippets{T1Bridge/E0Def}
\inputsnippets{T1Bridge/E0plusMultDef}

@@

\subsection{E0 as an ordinal notation}
The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
(see Chapter~\ref{chap:ON}).

@@

\subsection{A family of rapidly growing functions
(Library \texttt{GF\_alpha})}

The functions $F_\alpha$ are described in Section~\vref{sect:wainer}. The symbol \texttt{F\_} has type \texttt{E0 -> nat -> nat},
where \texttt{E0} is the type of well formed ordinal terms less than $\epsilon_0$.

\subsubsection{Importing the \texttt{F\_alpha} family}

We are now able to import \HydrasLib' $F_\alpha$ functions.

From~\href{../theories/html/gaia_hydras.G~F_alpha.html}%
{\texttt{gaia\_hydras.GF\_alpha}}.

@@

\begin{todo}
  Give equations for \texttt{F\_} (corresponding to the equation
  in \HydrasLib).
\end{todo}

@@

>>>>

\subsubsection{Remark}
\label{sect:gaia-masking}
We try to respect the following ``protocol''.
Let us assume we adapt some definitions and properties
from a given module \texttt{hydras.Epsilon0.Foo}, and a constant \texttt{bar} of this module, whose absolute name is
thus \texttt{hydras.Epsilon0.Foo.bar}.
\begin{enumerate}
\item We create a module \texttt{gaia\_hydras.GFoo}
  (``G'' for \gaia).
\item Inside \texttt{GFoo}, the identifier \texttt{bar}
  is reserved in priority to \gaia's version.
\HydrasLib' constructions are available through qualified names, like \texttt{Epsilon0.Foo.bar}, \texttt{Foo.bar}, or even a short notation like \texttt{hbar}  (``h'' for \HydrasLib).
\end{enumerate}

@@

Natural sum  (a.k.a. Hessenberg's  sum) is defined in
\href{../theories/html/hydras.Epsilon0.Hessenberg.html}{hydras.Epsilon0.Hessenberg} and is used for proving termination of all hydra battles (See Sect.~\ref{sect:hessenberg-def}).
We define this function in \href{../theories/html/gaia\_hydras.GHessenberg.html}{gaia\_hydras.GHessenberg} by delegation to \HydrasLib'.

\vspace{4pt} 
\noindent

@@

\inputsnippets{GHessenberg/oplusNf}
\inputsnippets{GHessenberg/oplusEquations}
\inputsnippets{GHessenberg/oplusEquationsb}
\inputsnippets{GHessenberg/oplusEquationsc}

@@

Like for Hessenberg's sum, we define a function \texttt{canon}
\emph{via} the
translations \texttt{g2h} and \texttt{h2g}.

\inputsnippets{GCanon/canonDef}
\inputsnippets{GCanon/g2hCanonRw}

@@

\subsubsection{Canonical sequences and the order \texttt{T1lt}}

\inputsnippets{GCanon/gcanonLt}
\inputsnippets{GCanon/gcanonLimitMono}
%\inputsnippets{GCanon/glimitCanonSNotZero}
\subsubsection{Canonical sequences and limits}

The following lemma (borrowed from\href{../theories/html/hydras.Epsilon0.Canon.html}%
{\texttt{hydras.Epsilon0.Canon}}) links limit ordinals to canonical sequences.

\inputsnippets{GCanon/gcanonLimitStrong}

@@

\inputsnippets{GCanon/gcanonLimitV2}
\inputsnippets{GCanon/gcanonLimitOf}

@@

 Paths with constant index and paths whose index is incremented by 1 at each step play an important role in the so-called Ketonen-Solovay machinery~\cite{KP81}.

 \inputsnippets{GPaths/pathsDefsb}

@@

\section{Rapidly growing arithmetic functions}
In this section, we consider a few families of arithmetic functions, indexed by ordinal numbers below $\epsilon_0$.
The type of such families is \texttt{E0 -> nat -> nat} where
\texttt{E0} is the type of ordinal terms in Cantor normal form, strictly below $\epsilon_0$.
In order to compare such functions, we define a few abstract properties in
\href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} (adapted from 
\href{../theories/html/hydras.Prelude.Iterates.html}{hydras.Prelude.Iterates}).

\inputsnippets{T1Bridge/MonoDef}

In \HydrasLib, these functions are defined by transfinite induction, using the \texttt{coq-equations} plug-in~\cite{sozeau:hal-01671777}. For that purpose, we defined a sigma-type
\texttt{hydras.Epsilon0.E0.E0} of ordinal-terms in Cantor normal form.

\subsection{A type for well-formed ordinal terms}
\label{sect:E0-gaia-hydras}
For compatibility's sake, we add a clone of \texttt{hydra-battles}' type \texttt{E0} defined in \href{../theories/html/hydras.Epsilon0.E0.html}{Epsilon0.E0}.

\inputsnippets{T1Bridge/E0Def}
\inputsnippets{T1Bridge/E0EqMixin, T1Bridge/E0EqType}

Many functions on type \texttt{T1} are ``lifted'' to \texttt{E0}.

\inputsnippets{T1Bridge/E0Defb}
\inputsnippets{T1Bridge/E0plusMultDef}
\inputsnippets{GCanon/E0CanonDef, GCanon/E0CanonLt}

@@

\subsubsection{E0 as an ordinal notation}
The module \href{../theories/html/gaia_hydras.T1Bridge.html}{gaia\_hydras.T1Bridge} defines an instance of class \texttt{ON E0lt compare}
(see Chapter~\ref{chap:ON}).

@@

\subsection{The \texorpdfstring{$H'_\alpha$}{H'\_alpha} family}

This variant of the Hardy hierarchy is described in Sect.~\vref{sect:hardy}, and adapted to \gaia in
Module \href{../theories/html/gaia_hydras.GHprime.html}{gaia\_hydras.GHprime}.

\inputsnippets{GHprime/Requirements, GHprime/Hprimedef}

Here is a list of lemmas proved by rewriting from \HydrasLib version ( ~\href{../theories/html/hydras.Epsilon0.Hprime.html}{Epsilon0.Hprime} ).

\subsubsection{Equations for $H'$}
In \HydrasLib', $H'_\alpha(i)$  is defined by transfinite induction over $\alpha$, using \texttt{coq-equations}. In the present module, equations for $H'$ are just imported as lemmas.

\inputsnippets{GHprime/HprimeEquations}

\subsubsection{Examples}

\inputsnippets{GHprime/Examples}
\inputsnippets{GHprime/Examplesb}

\subsubsection{Order and monotony properties}

\inputsnippets{GHprime/abstract}
\inputsnippets{GHprime/abstractb}
\inputsnippets{GHprime/abstractc}

\subsection{The \texorpdfstring{$F_\alpha$}{\texttt{F\_alpha}} hierarchy (Library \texttt{GF\_alpha})}

The functions $F_\alpha$ are described in Section~\vref{sect:wainer}.

%\subsubsection{Importing the \texttt{F\_alpha} family}

From~\href{../theories/html/gaia_hydras.GF_alpha.html}%
{\texttt{gaia\_hydras.GF\_alpha}}.

@@

The following equalities correspond to the equations of
~\href{../theories/html/hydras.Epsilon0.F_alpha.html}{Epsilon0.F\_alpha}

\inputsnippets{GF_alpha/FZeroEqn}
\inputsnippets{GF_alpha/FAlphaZeroEqn}
\inputsnippets{GF_alpha/FSuccEqn}
\inputsnippets{GF_alpha/FLimEqn}

@@

The $F_\alpha$ and $H'_\alpha$  hierarchies satisfy the inequality
$F_\alpha(n)\leq H'_{\omega^\alpha}(n)$ for any $\alpha$ and $n>0$.

\inputsnippets{GF_alpha/HprimeF}
